Testing Rate of Change Transition models

### 2023-05-18 minos/testing/rate_of_change_testing.Rmd
Rate of change models in a good working state for both income and SF_12_MCS now, success! Also added some testing plots to testing/rate_of_change_testing.Rmd

# Why wont this work with time as a factor?! Annoying!
```

@@

@@

ggplot(data = raw.quintile.sum.time, aes(x = time, y = diff.sd, group=quintile, color = quintile)) +
  geom_line()
```

@@

ggplot(data = raw.quintile.sum2, aes(x = income.quint, y = diff.sd, group=income.quint, fill = income.quint)) +
  geom_col()
```

@@

  return(raw)

  raw.time.quint <- raw %>%
    filter(time == 2009) %>%
    mutate(var.quint = cut(.data[[var]], breaks = unique(quantile(x = .data[[var]], prob = c(0:5)/5)), include.lowest = TRUE))# %>% , labels = c(1:5)
    #select(-var.mean)

@@

hh_income.stats.yr1 <- diff_by_year1_quintile(raw.dat, 'hh_income')
```

```{r}
test <- hh_income.stats.yr1

test2 <- test %>%
  filter(time == 2009)

print(sum(is.na(test2$hh_income)))
print(sum(test2$hh_income == -9))

print(min(test2$hh_income))
print(sum(test2$hh_income == 0))

ggplot(data = test2, aes(x = hh_income)) +
  geom_histogram()
```

>>>>

# Why wont this work with time as a factor?! Annoying!

rm(raw.sum)
```

@@

rm(raw.quintile.sum)
```

@@

ggplot(data = raw.quintile.sum.time, aes(x = time, y = diff.sd, group=quintile, color = quintile)) +
  geom_line()

rm(raw.quintile.sum.time)
```

@@

ggplot(data = raw.quintile.sum2, aes(x = income.quint, y = diff.sd, group=income.quint, fill = income.quint)) +
  geom_col()

rm(raw.time.quint, raw.quint2, raw.quintile.sum2)
```

@@

  raw.time.quint <- raw %>%
    filter(time == 2009) %>%
    mutate(varx = .data[[var]],
           var.quint = cut(varx, breaks = unique(quantile(x = varx, prob = c(0:5)/5)), include.lowest = TRUE))# %>% , labels = c(1:5)
    #select(-var.mean)

@@

#hh_income.stats.yr1 <- diff_by_year1_quintile(raw.dat, 'hh_income')
```

```{r}
#test <- hh_income.stats.yr1

#test2 <- test %>%
#  filter(time == 2009)

#print(sum(is.na(test2$hh_income)))
#print(sum(test2$hh_income == -9))

#print(min(test2$hh_income))
#print(sum(test2$hh_income == 0))

#ggplot(data = test2, aes(x = hh_income)) +
#  geom_histogram()
```

# Comparing SF12_MCS Diff

```{r}
cv.path <- 'Minos/data/final_US/cross_validation/simulation/'
cv.outpath <- 'Minos/output/cross_validation/'

raw.files <- list.files(cv.path, pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.cv <- do.call(rbind, lapply(raw.files, read.csv))

base.cv <- read_singular_local_out(cv.outpath, 'baseline')

## MODELS
mod.path <- 'Minos/data/transitions/cross_validation/'
sf12.mod <- readRDS(paste0(mod.path, 'SF_12/ols2/SF_12_2017_2018.rds'))
income.mod <- readRDS(paste0(mod.path, 'hh_income/ols2/hh_income_2019_2020.rds'))

## Model Summaries

```{r}
summary(sf12.mod)
```
```{r}
summary(income.mod)
```

```{r}
miss.values <- c(-9, -8, -7, -3, -2, -1, -10)
raw <- raw.cv %>%
  filter(!hh_income %in% miss.values, !SF_12 %in% miss.values) %>%
  group_by(pidp) %>%
  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time),
         diff.SF12_MCS = SF_12 - lag(SF_12, order_by = time)) %>%
  filter(time != 2009)

base <- base.cv %>%
  filter(!hh_income %in% miss.values, !SF_12 %in% miss.values) %>%
  group_by(pidp) %>%
  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time),
         diff.SF12_MCS = SF_12 - lag(SF_12, order_by = time)) %>%
  filter(time != 2014)
```

```{r}
raw.sub <- raw %>%
  select(pidp, time, diff.hh_income, diff.SF12_MCS) %>%
  mutate(source = 'raw')
base.sub <- base %>%
  select(pidp, time, diff.hh_income, diff.SF12_MCS) %>%
  mutate(source = 'output')

combined <- rbind(raw.sub, base.sub)

combined.sum <- combined %>%
  group_by(time, source) %>%
  summarise(hh_income = mean(diff.hh_income, na.rm = TRUE),
            SF12_MCS = mean(diff.SF12_MCS, na.rm = TRUE))

ggplot(combined.sum, aes(x = time, y = hh_income, group = source, color = source)) +
  geom_line()

ggplot(combined.sum, aes(x = time, y = SF12_MCS, group = source, color = source)) +
  geom_line()
```

```{r}
test <- base.cv %>%
  group_by(time) %>%
  summarise(mean.income = mean(hh_income, na.rm = TRUE),
            mean.SF12 = mean(SF_12, na.rm = TRUE))

ggplot(test, aes(x = time, y = mean.income)) +
  geom_line()

ggplot(test, aes(x = time, y = mean.SF12)) +
  geom_line()
```
